---
title: "STAT632 Final Project. Wine Quality"
author: "Chandnee & Jada"
date: "3/10/2022"
output: pdf_document
---

```{r}
#Libraries that will be used in the project.
library(tidyverse)
library(performance)
library(see)
library(patchwork)
library(MASS)
library(rpart)
```

```{r}
# Loading in the data, and ensuring we have the right data set.
vino <- read.csv("wine-quality-white-and-red.csv")
head(vino)
```

```{r}
# Confirming the dimensions of the data set.
  dim(vino)
  table(vino$type)
```

```{r}
vino_white <- vino %>%
  filter(type == "white") %>%
  drop_na()
```

## To better analyze, we will work onlye with the white wine data from the set.
## We have 4898 rows (Wines) by 13 columns(Aspects).
```{r}
dim(vino_white)
```

```{r}
View(vino_white)
```


```{r}
ggplot(vino_white, aes(x=alcohol, y=quality)) +
  geom_point()

ggplot(vino_white, aes(x=alcohol, y=factor(quality))) +
  geom_boxplot()

ggplot(vino_white, aes(x=volatile.acidity, y=quality)) +
  geom_point()

ggplot(vino_white, aes(x=volatile.acidity, y=factor(quality))) +
  geom_boxplot()
```
# Analysis
```{r}
# We will look at quality as a response variable to all of our predictor values.
lm_1 <- lm(quality ~ fixed.acidity + volatile.acidity + citric.acid
           + residual.sugar + free.sulfur.dioxide + total.sulfur.dioxide
           + density + pH + sulphates+alcohol,
           data = vino_white)
summary(lm_1)
```
## From summary of full model(lm_1), we get 8 predictor(fixed.acidity,volatile acidity,residual sugar,free.sulphur.dioxide,density,pH,sulphates,alcohol) which have most significant effect on the model. After that we will verify by using AIC.



```{r}
## For variable selection we are using AIC method here
step(lm_1)

```
## After using AIC, we can see that the last step of lm formula make with 8 same significant predictor like our summary method of lm_1.

```{r}
# fitting multiple linear regression by removing not significant predictor
lm_2 <- lm (quality ~ fixed.acidity + volatile.acidity + residual.sugar +free.sulfur.dioxide + density + pH + sulphates + alcohol, data = vino_white)

summary(lm_2)

```
## After removing some predictors, we can see from the reduced model(lm_2),all of the predictor has significant impact.
```{r}
# partial f test 
anova(lm_2, lm_1)

```
## From partial f test, we see that full model(lm1) p value is 0.7405>.01, so we fail to reject null hypothesis, and we can say that we can remove some predictors variable from full model to get better result.

```{r}
# After removing some predictor we see the correlation betwen remaining variables
pairs(quality ~ fixed.acidity + volatile.acidity + residual.sugar + 
    free.sulfur.dioxide + density + pH + sulphates + alcohol, data=vino_white)
```


```{r}
library(rpart)
t1 <- rpart(quality ~ ., data = vino_white)
```

```{r}
par(cex=0.7, xpd=NA)
plot(t1)
text(t1, use.n = TRUE)
```
## From tree model,Alcohol, Volatile acidity, density, free sulfur dioxide most impactful predictor in determining wine quality.

```{r}
par(mfrow=c(1,2), mar=c(4.5, 4.5, 2, 2))
plot(predict(lm_2), rstandard(lm_2),xlab="Fitted Values", ylab="Standardized Residuals")
abline(h=0)
qqnorm(resid(lm_2))
qqline(resid(lm_2))
hist(resid(lm_2))
```
## Standarized vs residual plot shows nonnormality due to nonconstant varience.From QQ plot we can see that normality of multiple linear regression not satisfied. Also from histogram of lm_2 looks like right skewed. We will do log transformation to get the normality

```{r}
library('performance')
library('see')
library('patchwork')
performance::check_model(lm_2)
```
```{r}
s1 <- summary(lm_1)
s2 <- summary(lm_2)
s1$adj.r.squared
s2$adj.r.squared
```




Adj. R-squared for lm_1:  0.2804
Adj. R-squared for lm_2:  0.2806
The Adj. R-squared only differs by a fourth decimal place. thus we see removing citric.acid and total.sulfur.dioxide do not impact our model.





```{r}
# Fitting model using log transformation
lm_3<-lm(log(quality) ~ log(fixed.acidity) + log(volatile.acidity) +log( residual.sugar) + log(
    free.sulfur.dioxide) + log(density) +log(pH) + log(sulphates) + log(alcohol),
                                                                         
     data=vino_white)
summary(lm_3)
```


```{r}
par(mfrow=c(1,2), mar=c(4.5, 4.5, 2, 2))
plot(predict(lm_2), rstandard(lm_2),xlab="Fitted Values", ylab="Standardized Residuals")
abline(h=0)
qqnorm(resid(lm_3))
qqline(resid(lm_3))
hist(resid(lm_3))
```
## After doing log transformation, we can see that histogram shows normal distribution but standardized vs residual plot still shows nonconstant varience and qq plot shows still some outliers.
```{r}
# Plot leverage
plot(hatvalues(lm_3), rstandard(lm_3),
xlab='Leverage', ylab='Standardized Residuals', xlim=c(0,0.35))
p<- 8
n<- nrow(vino_white)
abline(v = 2*(p+1)/n, lty=2)
abline(h = c(-4,4), lty=2)
```
```{r}
ind <- which(hatvalues(lm_3) > 0.01 & abs(rstandard(lm_3)) > 4)
vino_white[ind, ]
```
## fixed.acidity,volatile.acidity.citric acid has outliers.


